---
title: "flat_LPC_targets_project.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- 
Run this 'development' chunk

Store every call to library() that you need to run chunks line by line, as in a classical Rmd for analysis
-->

```{r development, include=FALSE}
library(testthat)
library(devtools)
library(usethis)
```

<!--
# Description of your package

This will fill the description of your package.
Fill and run the content of this chunk, before anything else. 

Note: when you will use other flat templates, this part will be in a separate file. Do not be surprised!
--> 

```{r description, eval=FALSE}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Miscellaneous Convenience Functions",
    Description = "A set of miscellaneous convenience functions.",
    `Authors@R` = c(
      person("Michael", "Levin", email = "mglev1n@gmail.com", role = c("aut", "cre"))
    )
  )
)
# Define License with use_*_license()
usethis::use_mit_license("Michael Levin")
```

## Introduction

The `targets` package is a tool for developing reproducible research workflows in R. Details of the package motivation and tools are described in detail: https://books.ropensci.org/targets/ and https://docs.ropensci.org/targets/.

### Example Workflow

1. Create a new R project

1. Run the `levinmisc::populate_targets_proj()` function in the R console - this function is described below in detail, and will initialize your project with helpful files/folders to start building your `targets` pipeline.

1. Modify the `Pipelines.qmd` file to specify your analyses. The `targets` documentation can be useful for specifics.

1. Knit/Render `Pipelines.qmd` to convert your markdown document into a series of R scripts that will actually run your analyses.

1. Run `submit-targets.sh` from a terminal window to submit your `targets` pipeline to the LPC for execution. Details of possible command line arguments to this script are described below.

1. Modify `Results.qmd` to present the results of your analyses. Rendering this file allows you to mix text describing your analyses/methods and include citations alongside the actual results of your pipeline. The `targets::tar_read()` function should be used heavily to load pre-computed results from your pipeline.

## populate_targets_proj()

The `populate_targets_proj` function can be run within a new project folder to initialize the project with the files/folders necessary for deploying a `targets` pipeline on the LPC at Penn. This includes creating LSF templates, a `Pipelines.qmd` file containing boilerplate for running analyses, and a `Results.qmd` file which can be used to visualize the results.

The `runtime` argument chooses the R installation the pipeline runs against. `"native"` (the default) loads the LPC's R module (`module load R/4.5`) and leaves the package library to `renv`. `"container"` instead runs the main `targets` process and every LSF worker inside the LPC's RStudio Singularity image, reading packages from a library built against that image's R, which is the right choice only for a project whose library was built against that image. Mixing the two fails at run time: a worker that loads a library built for a different R version reports errors such as `unused arguments (controller = ...)`, because the worker and the controller end up running different `crew` versions.

```{r function-populate_targets_proj}
#' Create a minimal targets template in the current project
#' 
#' @description
#' This function creates a minimal targets template in the current directory. This includes creating a `Pipelines.qmd` file containing boilerplate for running analyses, and a `Results.qmd` file which can be used to visualize the results. Parallelization of the pipeline is implemented using [targets::tar_make()] and `crew.cluster`, using pre-filled using parameters specific to the LPC system at Penn.
#'
#' Files are written to `path`, which defaults to the working directory. Passing it explicitly matters when the
#' project directory holds none of the markers a project search looks for (an `.Rproj` file, a `DESCRIPTION`,
#' a git repository, a `.vscode/` directory), because such a search otherwise walks up and writes into a parent
#' directory.
#'
#' The `runtime` argument selects the R installation the pipeline runs against, and is passed through to [use_crew_lsf()]. `"native"`, the default, loads the LPC's R module and leaves the package library to `renv`; `"container"` instead runs the main `targets` process and every worker inside the LPC's RStudio Singularity image.
#'
#' @param title (character) base name for project files (eg. "{title}-Pipeline.qmd" and "{title}-Results.qmd")
#' @param log_folder (character) directory for LSF logs
#' @param path (character) project directory the files are written to, defaulting to the working directory
#' @param runtime (character) R runtime the pipeline and its LSF workers use: `"native"` (the default) for the module-provided R, or `"container"` for the LPC RStudio Singularity image
#' @param overwrite (logical) overwrite existing template files
#'
#' @export
#' @concept targets

populate_targets_proj <- function(title,
                                  log_folder = "build_logs",
                                  path = ".",
                                  runtime = c("native", "container"),
                                  overwrite = FALSE) {
  runtime <- rlang::arg_match(runtime)

  # usethis::use_template() writes relative to the active usethis project, not
  # to the working directory, and usethis locates that project by walking *up*
  # from the working directory until a directory looks like one (.Rproj,
  # DESCRIPTION, .git, .vscode/, _quarto.yml, renv.lock, .here). A project
  # carrying none of those markers therefore sends every template into
  # whichever ancestor happens to have one. Pinning the active project to
  # `path` keeps the files where the user is working; force = TRUE skips that
  # same marker check, and local_project() restores the previous project when
  # this function exits.
  path <- fs::path_norm(fs::path_abs(path))
  fs::dir_create(path)
  usethis::local_project(path, force = TRUE, quiet = TRUE)

  # Create title
  if (missing(title)) {
    title <- basename(path)
  }

  # Check if files already exist
  if (file.exists("*-Results.qmd") | file.exists("*-Pipeline.qmd")) {
    cli::cli_alert_danger("Files already exist")
    overwrite <- yesno::yesno("Overwrite existing files?")
    if (!overwrite) cli::cli_abort("Exiting")
  }

  # Create build_logs folder
  fs::dir_create(path, log_folder)
  cli::cli_alert_success("Build log folder created at {.file {fs::path(path, log_folder)}}")

  # Copy batch script for job submission. The main {targets} process has to run
  # under the same R as the workers, so the runtime chooses the template.
  make_targets_template <- switch(runtime,
    container = ".make-targets.sh",
    native = ".make-targets-native.sh"
  )
  usethis::use_template(make_targets_template,
                        save_as = ".make-targets.sh",
                        package = "levinmisc")
  cli::cli_alert_success("Submission template created at {.file .make-targets.sh} using the {.val {runtime}} R runtime")

  # Copy batch script to submit targets job
  usethis::use_template("submit-targets.sh",
                        save_as = "submit-targets.sh",
                        package = "levinmisc")
  cli::cli_alert_success("Targets submission script created at {.file submit-targets.sh}")

  # Copy targets pipeline template
  usethis::use_template("Pipeline.qmd",
                        save_as = paste0(title, "-Pipeline.qmd"),
                        data = list(global_options = levinmisc::use_crew_lsf(runtime = runtime, path = path)),
                        package = "levinmisc")
  cli::cli_alert_success("Targets Pipeline template created at {.file {paste0(title, '-Pipeline.qmd')}}")

  # Copy targets results template
  usethis::use_template("Results.qmd",
                        save_as = paste0(title, "-Results.qmd"),
                        package = "levinmisc")
  cli::cli_alert_success("Targets Results template created at {.file {paste0(title, '-Results.qmd')}}")
}
```

<!--
Here is an example on how to use the function.
This should be a reproducible and working example
-->

```{r examples-populate_targets_proj, eval=FALSE}
#' \dontrun{
populate_targets_proj("test")
populate_targets_proj("test", runtime = "container")
#' }
```

The `populate_targets_proj()` function creates several files/folders within the project directory. `.make-targets.sh` and is a hidden helper file which is not designed for user interaction, but necessary for submission of jobs to the LSF scheduler; its contents follow the `runtime` argument, so that the main `targets` process runs under the same R as the workers. The other files are designed to be edited/used by the user:

  - `Pipeline.qmd` - Quarto markdown file which can be used to create a Target Markdown document that specifies a `targets` pipeline for your analyses. See: https://books.ropensci.org/targets/literate-programming.html#target-markdown for details. Remember to knit/render this document in order to generate the pipeline.
  
  - `Results.qmd` - Quarto markdown file which can be used to display the results generated by the `targets` pipeline specified in `Pipeline.qmd`
  
  - `build_logs/` - Directory where jobs logs are stored
  
  - `submit-targets.sh`- This is a bash script which can be used to run your `targets` pipeline, once `Pipeline.qmd` has been knit/rendered. This script can be run directly from the submission host.
  
## submit-targets.sh
This script is used to actually submit your pipeline to the LPC once `Pipeline.qmd` has been knit/rendered. This should be run from a terminal session from the root directory of your project. The only function of this script is to submit your pipeline to the LPC for analysis, and can be run directly from a submission host (eg. `scisub7`). The script can accept command line arguments, which can be useful for parallelizing your pipeline over multiple workers/CPUs:
  
```
Usage: ./submit-targets.sh [-n NUM_WORKERS] [-j JOB_NAME] [-o OUTPUT_LOG] [-e ERROR_LOG] [-q QUEUE] [-m MEMORY] [-s SLACK] [-h HELP]

Submit a job using the LSF scheduler with the specified number of CPUs and memory usage.

Options:
  -n NUM_WORKERS Number of workers (cpu cores) to request for running the targets pipeline (default: 1)
  -j JOB_NAME    Name of the job (default: make_targets)
  -o OUTPUT_LOG  Path to the output log file (default: build_logs/targets_%J.out)
  -e ERROR_LOG   Path to the error log file (default: build_logs/targets_%J.err)
  -q QUEUE       Name of the queue to submit the job to (default: voltron_normal)
  -m MEMORY      Memory usage for the job in megabytes (default: 16000)
  -s SLACK       Enable slack notifications; requires setup using slackr::slack_setup() (default: false)
  -h HELP        Display this help message and exit
```

### Slack Notifications

Slack can be used to automatically notify the user of pipeline start/finish using the `-s true` command line flag:

```
./submit-targets.sh -s true
```

Slack notifications are provided using the `slackr` package. The package must be configured separately before Slack notifications are enabled. See https://mrkaye97.github.io/slackr/index.html for more information about `slackr` setup and generation of an Slack API token.


<!--
Here are some unit tests to verify the function works as expected.
-->

```{r tests-populate_targets_proj}
test_that("populate_targets_proj works", {
  expect_true(inherits(populate_targets_proj, "function"))
})
# 
# # Functions to allow creating files in new temporary directory
# dir_empty <- function(x) {
#   unlink(x, recursive = TRUE, force = TRUE)
#   dir.create(x)
# }
# 
# test_with_dir <- function(desc, ...) {
#   new <- tempfile()
#   dir_empty(new)
#   withr::with_dir( # or local_dir()
#     new = new,
#     code = {
#       tmp <- capture.output(
#         testthat::test_that(desc = desc, ...)
#       )
#     }
#   )
#   invisible()
# }
# 
# test_with_dir("populate_targets_proj creates log folder", {
#   usethis::proj_set(force = TRUE)
#   populate_targets_proj()
#   expect_true(dir.exists("build_logs"))
# })
# 
# test_with_dir("populate_targets_proj creates .make-targets.sh", {
#   usethis::proj_set(force = TRUE)
#   populate_targets_proj()
#   expect_true(file.exists(".make-targets.sh"))
# })
# 
# test_with_dir("populate_targets_proj creates submit-targets.sh", {
#   usethis::proj_set(force = TRUE)
#   populate_targets_proj()
#   expect_true(file.exists("submit-targets.sh"))
# })
# 
# test_with_dir("populate_targets_proj creates pipeline and results .qmd templates", {
#   usethis::proj_set(force = TRUE)
#   populate_targets_proj()
#   expect_true(length(fs::dir_ls(".", glob = "*-Results.qmd")) > 0)
#   expect_true(length(fs::dir_ls(".", glob = "*-Pipeline.qmd")) > 0)
# })
```

### Use {crew} for parallelization

The `crew` and `crew.cluster` packages have enabled the use of heterogenous workers (<https://books.ropensci.org/targets/crew.html#heterogeneous-workers>), that can be used to deploy `targets` pipelines either locally or on HPC resources. The `use_crew_lsf()` function is designed to return a block of code to rapidly enable the use of heterogeneous workers on the Penn LPC. By default, this function creates workers that submit to different queues (eg. `voltron_normal`, `voltron_long`), and allocate different resources (eg. a "normal" worker will use 1 core and 16GB memory, while a "long" worker will use 1 core and 10GB memory).

`use_crew_lsf()` takes the same `runtime` argument as `populate_targets_proj()`, which controls the shell lines that open each worker script. Under `"container"` the worker exports `R_LIBS_USER` and `SINGULARITY_BIND` and then runs R through `singularity exec`; under `"native"` it runs `module load R/4.5` and nothing else, so the project library comes from `renv/activate.R` (`crew_options_lsf()` defaults `cwd` to the project directory). The two branches also differ in the `crew.cluster` interface they emit: the container branch keeps the older `lsf_*` arguments to `crew_controller_lsf()`, which is what the `crew.cluster` build inside the image understands, while the native branch uses the `crew_options_lsf()` interface that replaced them.
    
```{r function-use_crew_lsf}
#' Use crew lsf to execute a targets pipeline using the LSF HPC scheduler
#' 
#' @description
#' `r lifecycle::badge("experimental")`
#' 
#' This function returns a template for using `crew.cluster` in a targets project, enabling the parallel execution of a targets workflow. By default, the template is pre-filled using parameters specific to the LPC system at Penn. By default, this function creates workers that submit to different queues (eg. `voltron_normal`, `voltron_long`), and allocate different resources (eg. a "normal" worker will use 1 core and 16GB memory, while a "long" worker will use 1 core and 10GB memory).
#'
#' The `runtime` argument selects the R installation the workers run against. `"native"`, the default, loads the LPC's R module and leaves the package library to `renv`; `"container"` runs each worker inside the LPC's RStudio Singularity image instead. Use `"container"` only when the project library was built against that image: a worker that loads a library built for a different R version fails with errors such as `unused arguments (controller = ...)`, because the worker and the controller then run different `crew` versions.
#'
#' @param runtime (character) R runtime the LSF workers use: `"native"` (the default) for the module-provided R, or `"container"` for the LPC RStudio Singularity image
#' @param path (character) project directory the pipeline runs in, defaulting to the working directory. It supplies the controller names and, under the container runtime, the working directory each worker runs in
#'
#' @return A code block to copy/paste into a targets project
#'
#' @export
#' @concept targets

use_crew_lsf <- function(runtime = c("native", "container"), path = ".") {
  runtime <- rlang::arg_match(runtime)

  # The controller names and the container's working directory both come from
  # the project directory, taken as an argument rather than searched for: a
  # rprojroot-style search walks up from the working directory and lands in a
  # parent when the project carries no .Rproj, DESCRIPTION or .git marker.
  path <- fs::path_norm(fs::path_abs(path))
  title <- basename(path)

  # The "container" runtime executes each worker inside the LPC's RStudio
  # image, so `R_LIBS_USER` has to name a library built against that image's R.
  # The "native" runtime loads the LPC's R module instead and leaves the library
  # to `renv/activate.R`, because `crew_options_lsf()` defaults `cwd` to the
  # directory the controllers were built in - the project directory.
  container_image <- "/project/voltron/rstudio/containers/bioconductor-tidyverse_3.17.sif"
  container_library <- "$HOME/R/rocker-rstudio/bioconductor-tidyverse_3.17"
  container_bind <- "/project/:/project/, /appl/:/appl/, /lsf/:/lsf/, /scratch/:/scratch, /static:/static"
  r_module <- "R/4.5"

  # Shell lines that open each worker script, rendered as the elements of a
  # `script_lines = c(...)` call. crew appends its own `Rscript -e '...'` after
  # them, so under the container runtime the `singularity exec` line comes last
  # and ends in a line continuation.
  worker_lines <- function(queue, threads, indent) {
    lines <- c(
      paste0("#BSUB-q ", queue),
      if (runtime == "container") {
        paste0("export R_LIBS_USER=", container_library)
      } else {
        paste0("module load ", r_module)
      },
      "export TMPDIR=/scratch",
      paste0("export OMP_NUM_THREADS=", threads),
      if (runtime == "container") {
        c(
          paste0("export SINGULARITY_BIND='", container_bind, "'"),
          paste0("singularity exec --pwd ", path, " ", container_image, " \\")
        )
      }
    )

    paste0(strrep(" ", indent), encodeString(lines, quote = "\""), collapse = ",\n")
  }

  # A single `crew_controller_lsf()` call, rendered as the source the user
  # pastes into their pipeline. The native runtime uses the current
  # `crew_options_lsf()` interface; the container runtime keeps the `lsf_*`
  # arguments it replaced, which are what the crew.cluster build inside the
  # image understands.
  controller <- function(suffix, workers, queue, memory, threads = 1, cores = NULL) {
    if (runtime == "container") {
      glue::glue(
        "controller_lsf_{suffix} <- crew.cluster::crew_controller_lsf(
  name = '{title}_{suffix}',
  workers = {workers}L,{cores_line}
  lsf_memory_gigabytes_limit = {memory},
  script_dir = tools::R_user_dir('crew.cluster', which = 'cache'),
  lsf_log_output = 'build_logs/crew-%J.log',
  lsf_log_error = 'build_logs/crew-%J.err',
  script_lines = c(
{worker_lines(queue, threads, indent = 4)}
  ),
  verbose = TRUE
)",
        cores_line = if (is.null(cores)) "" else paste0("\n  lsf_cores = ", cores, ",")
      )
    } else {
      glue::glue(
        "controller_lsf_{suffix} <- crew.cluster::crew_controller_lsf(
  name = '{title}_{suffix}',
  workers = {workers}L,
  options_cluster = crew.cluster::crew_options_lsf(
    verbose = TRUE,
    script_directory = tools::R_user_dir('crew.cluster', which = 'cache'),
    script_lines = c(
{worker_lines(queue, threads, indent = 6)}
    ),
    log_output = 'build_logs/crew-%J.log',
    log_error = 'build_logs/crew-%J.err',
    memory_gigabytes_limit = {memory}{cores_line}
  )
)",
        cores_line = if (is.null(cores)) "" else paste0(",\n    cores = ", cores)
      )
    }
  }

  command <- glue::glue(
    "library(targets)
library(tarchetypes)
library(crew)
library(crew.cluster)

controller_local <- crew_controller_local(
  name = '{title}_local',
  workers = 1,
  seconds_idle = 10
)

{controller('normal', 20, 'voltron_normal', 16)}

{controller('long', 150, 'voltron_long', 10)}

{controller('highmem', 5, 'voltron_normal', 96)}

{controller('multicore', 5, 'voltron_normal', 96, threads = 16, cores = 16)}

# define some global options/functions common to all targets
options(tidyverse.quiet = TRUE)

tar_option_set(
  packages = c('tidyverse', 'targets', 'tarchetypes', 'arrow', 'vroom', 'data.table', 'gt', 'crew', 'crew.cluster'),
  controller = crew::crew_controller_group(controller_lsf_long, controller_lsf_normal, controller_lsf_highmem, controller_lsf_multicore, controller_local),
  resources = tar_resources(
    crew = tar_resources_crew(controller = '{title}_normal')
  ),
  storage = 'worker',
  retrieval = 'worker',
  cue = tar_cue(format = FALSE)
)"
  )

  cli::cli_h1("Copy the below code to the `global` chunk in your targets project:")

  return(command)
}
```

  
```{r example-use_crew_lsf, eval=FALSE}
#' \dontrun{
use_crew_lsf()
use_crew_lsf(runtime = "container")
#' }
```
  
```{r tests-use_crew_lsf}
test_that("use_crew_lsf works", {
  expect_true(inherits(use_crew_lsf, "function"))
})

test_that("use_crew_lsf defaults to the native runtime", {
  template <- suppressMessages(use_crew_lsf())
  expect_identical(
    as.character(template),
    as.character(suppressMessages(use_crew_lsf(runtime = "native")))
  )
  expect_true(grepl("module load R/", template, fixed = TRUE))
  expect_false(grepl("singularity exec", template, fixed = TRUE))
  expect_false(grepl("R_LIBS_USER", template, fixed = TRUE))
})

test_that("use_crew_lsf runs workers in the image under the container runtime", {
  template <- suppressMessages(use_crew_lsf(runtime = "container"))
  expect_true(grepl("singularity exec", template, fixed = TRUE))
  expect_true(grepl("export SINGULARITY_BIND=", template, fixed = TRUE))
  expect_false(grepl("module load", template, fixed = TRUE))
})

test_that("use_crew_lsf returns parseable R code for either runtime", {
  for (runtime in c("container", "native")) {
    template <- suppressMessages(use_crew_lsf(runtime = runtime))
    expect_true(length(parse(text = template)) > 0)
  }
})

test_that("use_crew_lsf rejects an unknown runtime", {
  expect_error(use_crew_lsf(runtime = "singularity"), "must be one of")
})
```
  

<!-- 
# Inflate your package

You're one inflate from paper to box.
Build your package from this very Rmd using `fusen::inflate()` 
-->


```{r development-inflate, eval=FALSE}
# Execute in the console directly
fusen::inflate(flat_file = "dev/LPC_targets_project.Rmd", vignette_name = "Create a New Targets Pipeline for LPC")
```

<!-- 
- Verify your `"DESCRIPTION"` file has been updated
- Verify your function is in `"R/"` directory
- Verify your test is in `"tests/testthat/"` directory
- Verify this Rmd appears in `"vignettes/"` directory 
-->
